# Data and demos

## Data in packages

Packages can also contain example data sets\sidenote{Full details are given at \url{http://goo.gl/Y4Srx4}.}. Data files\sidenote{Note that R code should be "self-
sufficient" and not make use of extra functionality provided by the package, so that the data file can also be used without having to load the package.} can be one of three types as indicated by their extension.

- plain R code (`.R` or `.r`)

- tables (`.tab`, `.txt` or `.csv`)

- `save()` images (`.RData` or `.rda`).

\marginnote{I tend to use RData files.}

\noindent{Data files live in the \texttt{data/} directory.}

\indent{Each data file should also have an associated help page. The easiest way to generate a help page is to use \textbf{roxygen2} and a dummy R function. Typically, I have a file called \texttt{data\_help\_files.R}, which has entries for the each data set. For example,}

```{r, eval=FALSE}
## Entry in data_help_files.R
## Name is name of the data set.
#' @name moby
#' @aliases moby_sample
#' @title Moby Dick word count
#' @description The frequency of occurrence of unique words
#' in the novel Moby Dick by Herman Melville.
#'
#' The data set moby_sample is 2000 values sampled from the
#' moby data set.
#' @docType data
#' @format A vector
#' @source M. E. J. Newman, "Power laws, Pareto distributions
#' and Zipf's law." Contemporary Physics 46, 323 (2005).
NULL
```

## Tasks

1. Create a `data/` directory.

2. Create the following data frame

   ```{r}
   example_data = data.frame(x = runif(10), y =        runif(10))
   ```
Now save\sidenote{Use the save function.} the data frame `example_data` in the `data/` directory
   ```{r, eval=FALSE}
   save(example_data,
   file= "data/example_data.RData")
   ```
3. Create a file called `data_help_files.R` in the `R/` directory and document your new data set.

4. Install and restart your package. Can you access the help page and the data set?

5. Check that your package still passes all tests\sidenote{CTRL + E.}.

## Demos

A demo is similar to function examples, but is typically longer and shows how to use multiple functions together. Demos are plain `.R` files that live in the `demo/` package directory. The demos are accessed with the `demo()` function.

\indent{In \texttt{demo/} directory, there should also be an \texttt{00Index} file, that lists
the demos\sidenote{There is a planned \texttt{demoTitle} tag for \textbf{roxygen2}, but currently this hasn’t been implemented.}. For example,\sidenote{Note the white space separation in
the \texttt{00Index}. Use at least four spaces
to avoid annoying error messages.}}

```
demo1   My very first demo
demo2   My very second demo
```

## Tasks

1. Create a `demo/` directory.

2. Create a file called `first.R` and save it in the `demo/` directory. In this file show how you can use some of your newly created function.

3. Add a `00Index` file to the `demo/` directory.

4. Build and check your package.
