# Documentation

Using `roxygen2` simplifies documentation\sidenote{The first package I wrote didn’t use \texttt{roxygen2} and it was a very painful experience.}. The premise of `roxygen2` is simple: describe your functions in comments next to their definitions and `roxygen2` will process your source code and comments to produce Rd files in the `man/` directory. In theory, you should never directly edit the `Rd` files.

## Tasks

1. Copy the following `roxygen2` descriptions to your `add` function

```{r}
#' @title A function for adding
#'
#' @description A really good adding function.
#' Perhaps the best function ever!
#'
#' A work of pure genius.
#' @param x a number
#' @param y another number
#' @return a number
#' @export
#' @examples
#' add(5, 10)
#' ## Can also use negative numbers
#' add(-5, 10)
add = function(x, y) {
if(!check_numeric(c(x, y))) stop("Not numeric")
x + y
}
```

2. Rebuild your package and look at the help file for the `add` function, i.e. `?add` . Run the examples via

\begin{table}[b]
  \centering
  \begin{tabular}{@{} ll}
    \toprule
    Tag name & Description \\
    \midrule
    \texttt{@title} & Short title for documentation page. \\
    \texttt{@description} & Longer description page. Skip a line for a \\
    &\qquad new paragraph.\\
    \texttt{@param} & Function parameter description. \\
    \texttt{@inheritParams} & Use the parameter definition from another function. \\
    \texttt{@export} & Add the function to the \texttt{NAMESPACE} file.\\
    \texttt{@return} & What does the function return, e.g. a data frame. \\
    \texttt{@examples} & Function examples (will be run when building). \\
    \texttt{@rdname} & Point multiple functions to the same help file,\\
    & \qquad e.g. \texttt{?substr}. \\
    \texttt{@seealso} & Pointers to other documentation pages. \\
    \texttt{@importFrom} & Import functions from other packages. \\
    \bottomrule
  \end{tabular}
  \caption{Useful \texttt{roxygen2} tags for documenting functions.}\label{T1.1}
\end{table}

```{r, eval=FALSE}
example(add)
```

3. Add a help page for the `subtract` function\sidenote{Install and restart.}.

4. Create a function called `multiply` and add an associated help page\sidenote{Install and restart.}.

5. Create a function called `times`

```{r}
times = function(x, y) multiply(x, y)
```

Export the `times` function\sidenote{Install and restart.}.

6. Use the `@rdname` tag above the `times` function to point to the `multiply` help page, i.e.

```{r}
#' @rdname multiply
```

Install and restart. Look at `?multiply` and `?times`. Now add \newline{\texttt{@examples} to the \texttt{times} function.} Look at the new `times` help \newline{page.}

## Importing functions

We often want to use functions from other R packages. When we do this we need to be explicit, i.e. state what we want and from where. The great thing about R packages, is that when we install a package, the dependencies are also automatically installed.

## Tasks

1. Install the package `jrPackage` . First, we need the `drat` package

```{r, eval=FALSE}
install.packages("drat")
```

Then we add the rcourses repo\sidenote{Run the \texttt{.libPaths()} function to see the repository location.}

```{r, eval=FALSE}
drat::addRepo("jr-packages")
```

Then install as usual

```{r, eval=FALSE}
install.packages("jrPackage")
```

2. The `jrPackage` contains a very useful function called `div` that we want to use

```{r}
library("jrPackage")
div(10, 2)
```

To use the div function within our package, we have to import it \newline{first}

```{r}
#' @importFrom jrPackage div
```

and add an entry to the `DESCRIPTION` file

```
     Imports: jrPackage
```

Create a function `divide` that uses the `div` function.

## The `DESCRIPTION` file

The `DESCRIPTION` file contains high level information about your package. For example, the package name, a brief description, the licence,
and your email address.

\indent{Open the \texttt{DESCRIPTION} file and update fields with relevant information. An example is given below.}

```
Package: pkg
Type: Package
Title: My First package
Version: 0.1
Date: 2016-11-01
Authors@R: person(given="Colin", family="Gillespie",
email="colin@jumpingrivers.com", role = c("aut", "cre"))
Maintainer: Colin Gillespie <colin@jumpingrivers.com>
Description: This is my very first package. It contains
exceedingly useful functions, such as add and subtract.
Make sure you add a couple of spaces to indent the
Description otherwise you will waste hours of your life
trying to find the bug.
License: GPL-2 | GPL-3
LazyData: TRUE
```

## Package checks

One of great things about R packages, is that there are a number of package checks that are available. These include

- Checking the syntax of the `DESCRIPTION` and `NAMESPACE` file.

- Checking your examples run.

- Checking all exported functions are documented.

## Tasks

1. Run the standard package checks on your package, via

```
      Build -> Check Package
```

Check that you package passes all tests\sidenote{CTRL + E.}. Fix any errors, warnings \newline{or notes.}

2. Add the following example to the `add` function

```{r}
#' add("A", "B")
```

Build the package. Does the package still build? Check the \newline{package. Does the package pass all tests?}