# Vignettes

If you want to include more extensive examples or even just further documentation, then you should consider creating a vignette:

\begin{quote}
  a \textbf{vignette} is a small illustration placed at the beginning or end of a book or chapter.
  
  \em \url{http://dictionary.reference.com/browse/vignette}
\end{quote}

We can view vignettes from other packages using the `vignette` \newline{function}

```{r}
vignette(package="knitr")
```

or to view in your web browser

```{r, , message=FALSE}
browseVignettes(package="knitr")
```

## Example: markdown vignettes

Vignettes are stored in the `vignettes/` directory. The simplest vignette uses R-markdown and is formatted by the `knitr` package. To create a package vignette, we simply place the file in the `vignettes/` directory. 

\marginnote{The output style is \texttt{html\_vignette}. This is more lightweight (in terms of file size) than the standard \texttt{html\_document}}

\indent{For example, suppose we have a file \texttt{intro.Rmd} that contains the following text:}

\marginnote{The UTF-8 line specifies the file encod-
ing.}

```{r, eval=FALSE}
---
title: "My very first vignette"
author: "Colin Gillespie and Robin Lovelace"
output: rmarkdown::html_vignette
vignette: >
   %\VignetteIndexEntry{My very first vignette}
   %\VignetteEngine{knitr::rmarkdown}
   %\VignetteEncoding{UTF-8}
---

## My first package

This is my **first** package vignette. I can
include mathematics, such as $x^2$. R code is also
nicely formatted and displayed.

# ```{r}
# x = runif(10)
# ```
and plots
# ```{r}
# plot(x)
# ```
```

1. The first few lines register `knitr` as the vignette engine and provide an entry for the list of vignettes.

2. We have used markdown to add simple styling\sidenote{See \url{http://goo.gl/y9FO5} and the RStudio markdown reference sheet for further commands.}. For example \texttt{**first**} becomes **first** and \texttt{\$x\textasciicircum 2\$} becomes $x^2$.

3. R code is executed in the ``` regions.

## Tasks

1. Create a `vignettes/` directory in your package.

2. Create an R markdown file

```
   File -> New File -> R markdown
```

and save it in the `vignettes/` directory.

3. Copy the markdown example above into your file and `knit` that file.

4. Add

```
   Suggests: knitr
   VignetteBuilder: knitr
```

to the DESCRIPTION file\sidenote{Notice the \texttt{knitr} is only a \textbf{suggested} package, since it’s not essential.}. This will tell your package to build \newline{vignettes using \texttt{knitr}.}

5. Vignettes won’t actually be built unless you are creating a source bundle\sidenote{If you submit your package to CRAN, then the vignettes will be built ready for distribution.}. To install a package with vignettes included, we first create the source package

```
   Build -> Build Source Package
```

and then install the package from source

```{r, eval=FALSE}
install.packages("pkg_1.0.tar.gz", repos=NULL,
type="source")
```

Build and install your package. Check that you can access the \newline{vignette.}

6. Check and ensure that your package passes all tests\sidenote{CTRL + E.}.

## Package level documentation

You package typically also has documentation associated with the package name. For example,

```{r, eval=FALSE, tidy=FALSE}
library("pkg")
?pkg
```

\noindent{should bring up an overview of your package. Again we use \texttt{roxygen2} to generate the man page. Go to}

\begin{center}
  \url{http://goo.gl/W2tJrF}
\end{center}
to view the entry for the `poweRlaw` package.

## Tasks

1. Create a file called `pkg-package.R` in the `R/` directory.

2. Using the `poweRlaw` package as an example, create a man page for your package\sidenote{What do you think the \texttt{@aliases} tag does?}.

3. Do one final check and ensure that your package passes all tests.